Hydra Battles and Cie

### 2021-07-26 doc/chapter-primrec.tex
more snippets

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}}.
\index{hydras}{Library Prelude!iterate}
\begin{Coqsrc}
Fixpoint iterate {A:Type}(f : A -> A) (n: nat)(x:A) :=
  match n with
  | 0 => x
  | S p => f (iterate  f p x)
  end.
\end{Coqsrc}

\begin{Coqsrc}
Lemma iterate_le_n_Sn:
forall f : nat -> nat,
(forall x : nat, x <= f x) ->
forall n x : nat, iterate f n x <= iterate f (S n) x.
\end{Coqsrc}

@@

\begin{exercise}
The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions\footnote{post on \href{https://stackoverflow.com/questions/10292421/error-in-defining-ackermann-in-coq} by Anton Trunov.} of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
Prove that the four functions are extensionnally equal.
\end{exercise}

@@

\begin{Coqsrc}
Lemma Ack_0 : Ack 0 = S.
Proof refl_equal.

Lemma Ack_S_0 m : Ack (S m) 0 = Ack m 1. 
Proof. reflexivity. Qed.

Lemma Ack_S_S : forall m p,
    Ack (S m) (S p) = Ack m (Ack (S m) p).
Proof. reflexivity. Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma Ack_1_n n: Ack 1 n = S  (S n).

Lemma Ack_2_n n: Ack 2 n = 2 * n + 3.

Lemma Ack_3_n n: Ack 3 n = exp2 (S (S (S n))) - 3.

Lemma Ack_4_n n: Ack 4 n = hyper_exp2 (S (S (S n))) - 3.
\end{Coqsrc}

@@

\noindent
\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}}.
\begin{Coqsrc}
Lemma nested_Ack_bound : forall k m n, 
    Ack k (Ack m n) <= Ack (2 + max k m) n.
\end{Coqsrc}

@@

\noindent
\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.
\begin{Coqsrc}
Theorem Ackn_IsPR (n: nat) : isPR 1 (Ack n).
\end{Coqsrc}

@@

\index{coq}{Commands!Scheme}
\begin{Coqsrc}
Scheme PrimRec_PrimRecs_ind := Induction for PrimRec
  Sort Prop
  with PrimRecs_PrimRec_ind := Induction for PrimRecs 
  Sort Prop.
\end{Coqsrc}

\begin{Coqanswer}
PrimRec_PrimRecs_ind :
forall (P : forall n : nat, PrimRec n -> Prop)
  (P0 : forall n n0 : nat, PrimRecs n n0 -> Prop),
(* successor *)
P 1 succFunc ->

(* zero *)
P 0 zeroFunc ->

(* projections *)
(forall (n m : nat) (l : m < n), P n (projFunc n m l)) ->

(* composition *) 
(forall (n m : nat) (g : PrimRecs n m),
      P0 n m g -> forall h : PrimRec m, P m h -> 
      P n (composeFunc n m g h)) ->

(* primitive recursion *)
(forall (n : nat) (g : PrimRec n),
 P n g ->
    forall h : PrimRec (S (S n)), P (S (S n)) h -> 
     P (S n) (primRecFunc n g h)) ->

(* empty list of functions *)
(forall n : nat, P0 n 0 (PRnil n)) ->

(* add a function to a list *)
(forall (n m : nat) (p : PrimRec n),
   P n p -> 
   forall p0 : PrimRecs n m, P0 n m p0 -> 
   P0 n (S m) (PRcons n m p p0)) ->

(* conclusion ! *)
forall (n : nat) (p : PrimRec n), P n p
\end{Coqanswer}

For instance, proving a property shared by any primitive recursive function of arity 2 leads to consider the case where that function is obtained by composition with a function of any 
arity $m$. The same problem happens with primitive 
recursion, where a function of arity $n$ is built out of a function of arity $n+1$ and a function of arity $n-1$.

@@

\index{coq}{Dependently typed functions}
\begin{Coqsrc}
Fixpoint max_v {n:nat} : forall (v: Vector.t nat n) , nat :=
  match n as n0 return (Vector.t nat n0 -> nat)
  with
    0 => fun v => 0
  | S p => fun (v : Vector.t nat (S p)) =>
             max (Vector.hd v) (max_v  (Vector.tl v))
  end.

Lemma max_v_2 : forall x y,  max_v (x::y::nil) = max x y.

Lemma max_v_lub : forall n (v: t nat n) y,
    (Forall (fun x =>  x <= y) v) -> max_v v <= y.

Lemma max_v_ge : forall n (v: t nat n) y,  In  y  v -> y <= max_v v.
\end{Coqsrc}

@@

\begin{Coqsrc}
Fixpoint evalList (m : nat) (l : Vector.t nat m) {struct l} :
 naryFunc m -> nat :=
  match l in (Vector.t _ m) return (naryFunc m -> nat) with
  | Vector.nil => fun x : naryFunc 0 => x
  | Vector.cons a n l' => fun x : naryFunc (S n) => evalList n l' (x a)
  end.
\end{Coqsrc}

@@

\index{coq}{Dependently typed functions}
\begin{Coqsrc}
(**  uncurried apply:

[v_apply f (x1::x2:: ... ::xn::nil)]  is [f x1 x2 ... xn] 
 *)

Notation "'v_apply' f v" := (evalList _ v f) 
     (at level 10, f at level 9).

Example Ex2: forall (f: naryFunc 2) x y,
    v_apply f (x::y::nil) = f x y.
Proof.   intros; now cbn. Qed.

Example Ex4: forall (f: naryFunc 4) x y z t,
    v_apply f (x::y::z::t::nil) = f x y z t.
Proof.  intros; now cbn. Qed.
\end{Coqsrc}

@@

\index{coq}{Dependently typed functions}
\begin{Coqsrc}
Definition majorized {n} (f: naryFunc n) (A: naryFunc 2) : Prop :=
  exists (q:nat), forall (v: t nat n),
      v_apply f v <= A q  (max_v v).

Definition majorizedPR {n} (x: PrimRec n) A := 
           majorized (evalPrimRec n x) A.

(** For vectors of functions *)

Definition majorizedS {n m} (fs : Vector.t (naryFunc n) m)
           (A : naryFunc 2):=
  exists N, forall (v: t nat n),
      max_v (map (fun f => v_apply f v) fs) <= A N (max_v v).

Definition majorizedSPR {n m} (x : PrimRecs n m) :=
  majorizedS (evalPrimRecs _ _ x).
\end{Coqsrc}

>>>>

\vspace{4pt}
\noindent
\emph{From \href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}}.
\index{hydras}{Library Prelude!iterate}

\input{movies/snippets/Iterates/iterateDef}

\input{movies/snippets/Iterates/iterateLeNSN}

@@

\begin{exercise}
The file \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack} presents two other definitions of the Ackermann functions based on the lexicographic ordering on $\mathbb{N}\times\mathbb{N}$.
Prove that the four functions are extensionnally equal.
\end{exercise}

@@

\input{movies/snippets/Ack/AckRewrite}

@@

\input{movies/snippets/Ack/Ack1N}
\input{movies/snippets/Ack/Ack2N}
\input{movies/snippets/Ack/Ack3N}
\input{movies/snippets/Ack/Ack4N}

@@

\noindent
\emph{From \href{../theories/html/hydras.MoreAck.Ack.html}{MoreAck.Ack}}.

\input{movies/snippets/Ack/nestedAckBound}

@@

\noindent

\emph{From \href{../theories/html/hydras.MoreAck.AckNotPR.html}{MoreAck.AckNotPR}}.

\input{movies/snippets/AckNotPR/AckNIsPR}

@@

\index{coq}{Commands!Scheme}

\input{movies/snippets/primRec/SchemePrimRecInd}

Please note that, in order to prove a property shared by any primitive recursive function of, say, arity 2, this induction scheme  leads you to consider an expension of the considered property to primitive recursive function of any arity.

@@

\index{coq}{Dependently typed functions}

\input{movies/snippets/MoreVectors/maxvDef}

\input{movies/snippets/MoreVectors/maxvLemmasa}

\input{movies/snippets/MoreVectors/maxvLemmasb}

\input{movies/snippets/MoreVectors/maxvLemmasc}

@@

\input{movies/snippets/primRec/evalListDef}

@@

\index{coq}{Dependently typed functions}

\input{movies/snippets/AckNotPR/vApply}

@@

\index{coq}{Dependently typed functions}

\input{movies/snippets/AckNotPR/majorizedDefs}
